Towards ....

### 2015-03-24 paper/nodalida2015.tex
more text

\documentclass[11pt]{article}
\usepackage{nodalida2015}
\usepackage{times}
\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{latexsym}
\usepackage{listings}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

@@

\author{Inari Listenmaa \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt inari@chalmers.se} \\\And
  Koen Claessen \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt koen@chalmers.se} \\}

@@

\cite{lager1998, lager2000}  CG and FSIG (Finite State Intersectional Grammar) inspired POS tagging using theorem prover.
% What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.

@@

\section{Implementation using SAT}
\label{sect:pdf}

@@

SAT solving is used for many applications where constraints must be
satisfied, such as software and hardware verification.
Application of logic to POS tagging or shallow parsing isn't new;
Lager~\shortcite{lager1998, lager2000} presents POS tagging rules as logic, and uses
a theorem prover (DisLog, TODO cite) to implement small constraint rules.
Inspired by the approach, our implementation tries to emulate the CG
formalism in behaviour and features offered.

We make each analysis a variable. For instance, the fifth line denotes
a hypothesis that the correct analysis of the 3rd item in the sentence
is \texttt{<V,Sg>}. We call this \texttt{v5}.

\begin{lstlisting}
((1,["<the>","the",det]),v0)
((2,["<bear>","bear",n,sg]),v1)
((2,["<bear>","bear",vblex,pl]),v2)
((2,["<bear>","bear",vblex,fin]),v3)
((3,["<sleeps>","sleep",n,pl]),v4)
((3,["<sleeps>","sleep",vblex,sg,p3]),v5)
\end{lstlisting}

@@

>>>>

\documentclass[11pt]{article}
\usepackage{nodalida2015}
\usepackage{times}
\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{latexsym}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

@@

\author{Inari Listenmaa \and Koen Claessen \\
  Chalmers University of Technology, Gothenburg, Sweden \\
  {\tt \{inari,koen\}@chalmers.se} }

@@

Application of logic to POS tagging or shallow parsing isn't new;
\cite{lager1998, lager2000} present constraint rules as 
a disjunctive logic program.

Our implementation is inspired by this approach,
but unlike Lager's our implementation tries to emulate the CG
formalism in behaviour and features offered.
% What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.

@@

\section{Implementation using SAT solver}
\label{sect:pdf}

@@

SAT solving is used for many applications where constraints must be
satisfied, such as software and hardware verification.

Let us demonstrate the approach with the example sentence 
\emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:

\begin{verbatim}
"<the>"
        "the" det
"<bear>"
        "bear" n sg
        "bear" v pl
"<sleeps>"
        "sleep" n pl
        "sleep" v sg p3
\end{verbatim}

We make each analysis a variable \texttt{v1}--\texttt{v5}. 
For instance, \texttt{v3} is a hypothesis that the correct analysis 
for the word form ``bear'' is \texttt{"bear" n pl}.
Before applying rules, we must make sure that in each position
at least one of the analyses is true. We add our first disjunctions:

\begin{itemize}
\item [] $v1$ \\ $v2 \vee v3$ \\ $v4 \vee v5$
\end{itemize}

The first

@@

A typical CG rule sequence starts with stronger rules
